
```{r Load, include=FALSE}
library(Ecdat)
require(FinTS)
library(forecast)
data(IncomeUK)
head(IncomeUK)
consumption = IncomeUK[,2] 

```

#Problem 1 Describe the behavior of consumption. What types of differencing, seasonal, nonseasonal, or both, would you recommend? Do you recommend fitting a seasonal ARIMA model to the data with or without a log transformation? Consider also using ACF plots to help answer these questions.

```{r Q1}
plot(consumption)
acf(consumption)
plot(diff(consumption))
acf(diff(consumption))
logConsumption=log(consumption) #consumption is a time series
plot(logConsumption)
acf(logConsumption)
```

// From the first plot of consumption we can see that it is not a stationary time series. The graph increases as the time progressesing therefore it does not have a standard mean.
//The ACF tests confirms what was said earlier about the time series. The ACF plot getting closer to 0.
//Plotting the diff(consumption), we can see that the time series seems more stationary that before and we reconfirm this with the ACF plot. //This acf plot of the diff(consumption) plot shows us that the time series occurs periodically, hence suggesting seasonal effect.
//It is preferable to use both seasonal and non seasonal differencing. 
//The acf plot of log(consumption) still suggests that series is non stationary. Therefore producing no differnce. 
//A seasonal ARIMA model without log transform would be ideal. 

#Problem 2 Regardless of your answers to Problem 1, find an ARIMA model that provides a good fit to log(consumption). What order model did you select? (Give the orders of the nonseasonal and seasonal components.)
```{r Q2, echo=FALSE}
plot(pacf(log(consumption),lag.max=16))
m1=auto.arima(logConsumption, ic='aic')
summary(m1)

```
//Series: logConsumption 
ARIMA(0,1,0)(0,1,1)[4]
Non seasonal: (p,d,q)=(0,1,0)
Seasonal: (P,D,Q) = (0,1,1)
Seasonal period: 4

#Problem 3 Check the ACF of the residuals from the model you selected in Problem 2. Do you see any residual autocorrelation?
```{r Q3, echo=FALSE}
acf(m1$residuals)
Box.test(m1$residuals, lag=16,type="Ljung")

```
//The Box-Ljung test result show that p>0.01 and we fail to reject the null hypothesis that there is no residual auto correlation. 

#Problem 4 Apply auto.arima to log(consumption) using BIC. What model is selected?
```{r Q4, echo=FALSE}
m2=auto.arima(logConsumption, ic='bic')
summary(m2)
```
Series: logConsumption 
ARIMA(0,1,0)(0,1,1)[4]
//The same model as the above question is selected for both AIC and BIC models.
//(1-B)*(1-B^4)*logConsumption$mean(t)=theta(0) +(1-theta(4) *(B^4))epsi(t)
//(1-B)*(1-B^4)*logConsumption$mean(t)=(1-0.5348 * (B^4))epsi(t)


#Problem 5 Forecast log(consumption) for the next eight quarters using the models you found in Problems 2 and 4. Plot the two sets of forecasts in side-by- side plots with the same limits on the x- and y-axes. Describe any differences between the two sets of forecasts. Using the backshift operator, write the models you found in problems 2 and 4.
```{r Q5, echo=FALSE}
foreAutoArima = forecast(m1,h=8)
foreAutoArima_log = forecast(m2,h=8)
par(mfrow=c(1,2))
plot(foreAutoArima$mean,xlim=c(1985.5,1987.25),ylim=c(10.86,11))
plot(foreAutoArima_log$mean,xlim=c(1985.5,1987.25),ylim=c(10.86,11))
```
//There are no noticable differences between the two plots. 

#Problem 6 Include the variable include log(Income) as an exogenous variable to forecast log(consumption) using auto.arima.  According to the AIC, is this model better than the previous models? (Hint: use xreg to include exogenous variables in arima and auto.arima)
```{r Q6, echo=FALSE}
Income=IncomeUK[,1]
Income_model=auto.arima(logConsumption,xreg=log(Income))
summary(Income_model)
```
Series: logConsumption 
Regression with ARIMA(2,0,0)(0,1,1)[4]
//According to the AIC of -278.4, this is better than the previous AIC of -275.55. 