---
title: "Lab2 Volatility BIA656A"
author: "Rana Putta"
date: "10/14/2020"
output: html_document
---

1. Consider the daily returns of the exchange trade fund (ETF) SPDR S&P 500 of State
Street Global Advisors from September 4, 2001 to September 30, 2011. The tick symbol
is SPY and there are 2535 observations. The simple returns are available from CRSP
and in the file d-spy-0111.txt. Transform the simple returns to log returns.

```{r Q1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fGarch)
require(FinTS)
require(rugarch)
da=read.table("d-spy-0111.txt",header=T)  #<== Load data with header
spy=log(da$rtn+1)
acf(spy)
t.test(spy)
ArchTest(spy)  

m1=garchFit(~arma(1,1)+garch(2,1),data=spy,trace=F,cond.dist="sstd") 
summary(m1)
sresi=m1@residuals/m1@sigma.t 
acf(sresi)
Box.test(sresi,lag=15,type="Ljung")
Box.test(sresi^2,lag=15,type="Ljung")
qqnorm(sresi)


m2=garchFit(~arma(0,0)+garch(2,1),data=spy,trace=F,cond.dist="std")
#summary(m2) 

m3=garchFit(~arma(0,1)+garch(2,1),data=spy,trace=F,cond.dist="sstd")
#summary(m3)

m4=garchFit(~arma(1,0)+garch(2,1),data=spy,trace=F,cond.dist="sstd")
#summary(m4) 


m1@fit$ics[1]
m2@fit$ics[1]
m3@fit$ics[1]  
m4@fit$ics[1]

sresi2=m3@residuals/m3@sigma.t  ## For model checking
acf(sresi2,lag.max = 10)
Box.test(sresi2,lag=10,type="Ljung")
Box.test(sresi2^2,lag=10,type="Ljung")


spec1=ugarchspec(variance.model=list(model="apARCH",garchOrder=c(2,1)),
                 mean.model=list(armaOrder=c(0,0)),distribution.model="std")

m5=ugarchfit(spec=spec1,data=spy)
m5  ### see output


### prediction, 1-step to 5-step ahead
p1 <- ugarchforecast(m5,n.ahead=5)
p1

sigma(p1) ### volatility prediction

```
(a) Is the expected log return zero? Are there any serial correlations in the log
returns? Is there ARCH effect in the log returns?
//
No the expected log return is not zero. The p-value from thee t-test is 0.7909 and we cannot reject the null hypothesis. 
Yes, there is ARCH effect in the log returns.

(b) Fit a Gaussian ARMA-GARCH model for the log return series. Perform model
checking, obtain the QQ-plot of the standardized residuals, and write down the
fitted model. [Hint: Try GARCH(2,1).]
//
garchFit(~arma(1,1)+garch(2,1),data=spy,trace=F,cond.dist="sstd")
AIC -6.340424 

(c) Build an ARMA-GARCH model with Student-t innovations for the log return
series. Perform model checking and write down the fitted model.
//
m1=garchFit(~arma(1,1)+garch(2,1),data=spy,trace=F,cond.dist="sstd")

(d) Fit an ARMA-APARCH model with Student-t innovations to the data. Write
down the fitted model and perform 1- to 5-step ahead predictions of the series
and its volatility.
//
spec1=ugarchspec(variance.model=list(model="apARCH",garchOrder=c(2,1)),
                 mean.model=list(armaOrder=c(0,0)),distribution.model="std")

2. Consider the monthly stock returns of the Coca-Cola Company (KO) from January
1961 to September 2011. The simple returns are available from CRSP and in the file
m-ko-6111.txt. Transform the simple returns to log returns.
```{r Q2, echo = FALSE}
library(fGarch)
require(FinTS)
require(rugarch)

da1=read.table("m-ko-6111.txt",header=T)  #<== Load data with header
ko=log(da$rtn+1)
acf(ko)
t.test(ko)
ArchTest(ko)

m1=garchFit(~garch(2,1),data=ko,trace=F,cond.dist="sstd")
summary(m1) 
sresi=m1@residuals/m1@sigma.t  ## For model checking
acf(sresi)
Box.test(sresi,lag=15,type="Ljung")
Box.test(sresi^2,lag=15,type="Ljung")

m2=garchFit(~garch(1,1),data=ko,trace=F,cond.dist="std")
summary(m2) 
sresi1=m2@residuals/m2@sigma.t  ## For model checking
acf(sresi)
Box.test(sresi1,lag=15,type="Ljung")
Box.test(sresi1^2,lag=15,type="Ljung")

m1@fit$ics[1]
m2@fit$ics[1]
qqnorm(sresi)

spec1=ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(2,1)),
                 mean.model=list(armaOrder=c(0,0)))
m7=ugarchfit(spec=spec1,data=ko)
m7  ### see output

p1 <- ugarchforecast(m7,n.ahead=5)
p1
sigma(p1)
```
(a) Is the expected monthly log return zero? Is there any serial correlation in the log
returns? Is there any ARCH effect in the log returns?
//
Yes there is ARCH effect in the log 

(b) Build a Gaussian GARCH model for the log returns. Perform model checking
and write down the fitted model.
//
m1=garchFit(~garch(2,1),data=ko,trace=F,cond.dist="sstd")

(c) Build a GARCH model with Student-t innovations for the log returns. Perform
model checking, obtain the QQ-plot of the standardized residuals, and write down
the fitted model. Also, obtain 1- to 5-step ahead volatility predictions.
//
spec1=ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(2,1)),
                 mean.model=list(armaOrder=c(0,0)))
                 
3. Consider again the monthly log returns of KO stock. Multiple the log returns by 100.
i.e., use percentage log returns
```{r Q3, echo = FALSE}
library(fGarch)
require(FinTS)
require(rugarch)

da1=read.table("m-ko-6111.txt",header=T)  #<== Load data with header
ko=log(da$rtn+1)*100
acf(ko)
t.test(ko)

spec3=ugarchspec(variance.model=list(model="tGARCH"),mean.model=list(armaOrder=c(0,0)))
k1=ugarchfit(spec=spec3,data=ko)
k1

variance.model=list(model="sGARCH",garchOrder=c(1,1), submodel="GARCH")
mean.model=list(armaOrder=c(1,0),include.mean=TRUE,archm=T,archpow=2)
spec4=ugarchspec(variance.model=variance.model,mean.model=mean.model,distribution.model="norm")

k2=ugarchfit(spec=spec4,data=ko)
k2

```
(a) Fit a TGARCH model to the series. Perform model checking and write down the
fitted model. Is the leverage effect different from zero?
//
spec3=ugarchspec(variance.model=list(model="tGARCH"),mean.model=list(armaOrder=c(0,0)))
Leverage effect is different from zero.


(b) Fit a GARCH-M model to the series. Perform model checking and write down
the fitted model. Is the risk premium significant? Why?
//
variance.model=list(model="sGARCH",garchOrder=c(1,1), submodel="GARCH")
mean.model=list(armaOrder=c(1,0),include.mean=TRUE,archm=T,archpow=2)
spec4=ugarchspec(variance.model=variance.model,mean.model=mean.model,distribution.model="norm")
Risk premium is significant
